### 2017-10-09 Proposal/paper.tex

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage[normalem]{ulem}
% Copyright
\setcopyright{none}

\input{annotations}

\begin{document}

\title{TCPTokens: Introducing Currency into Data Center Congestion Control}

\author{Anand Jayarajan}
\affiliation{%
  \institution{University of British Columbia}
}
\email{anandj@cs.ubc.ca}

\author{Michael Przystupa}
\affiliation{%
  \institution{University of British Columbia}
}
\email{michael.przystupa@gmail.com}

\author{Robert Reiss}
\affiliation{%
  \institution{University of British Columbia}
}
\email{rreiss@cs.ubc.ca}

\author{Fabian Ruffy}
\affiliation{%
  \institution{University of British Columbia}
}
\email{fruffy@cs.ubc.ca}

\begin{abstract}

\end{abstract}

\keywords{TCP, Congestion Control, SDN, Data center}

\maketitle

\input{intro}

\input{relatedworks}
\input{design}
%\input{motivation}
\input{implementation}
\input{eval}

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}