### 2017-10-10 Proposal/paper.tex

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage[normalem]{ulem}
% Copyright
\setcopyright{none}

\input{annotations}

\begin{document}

\title{TCPTokens: Introducing Currency into Data Center Congestion Control}

\author{Anand Jayarajan}
\affiliation{%
  \institution{University of British Columbia}
}
\email{anandj@cs.ubc.ca}

\author{Michael Przystupa}
\affiliation{%
  \institution{University of British Columbia}
}
\email{michael.przystupa@gmail.com}

\author{Robert Reiss}
\affiliation{%
  \institution{University of British Columbia}
}
\email{rreiss@cs.ubc.ca}

\author{Fabian Ruffy}
\affiliation{%
  \institution{University of British Columbia}
}
\email{fruffy@cs.ubc.ca}

\begin{abstract}
Datacenter networks have become a hotbed for research in the recent past. Many works have been published on improving the bisection bandwidth, reducing cost, and improving congestion control as well as latency. As a core contribution of this research, centralized scheduling has entered the stage as dominant strategy to manage flows. Benefitting from a global view and full control over the network, centralized schedulers are able to enforce fine-grained traffic control, frequently achieving near-optimal bandwidth optimization. However, modern congestion control algorithms and schedulers are still fundamentally reactive. Many algorithms are designed to only respond to packet loss or significant increase in latency, not to actively prevent it.
In this project, we plan to explore the opportunities of employing a preemptive and tightly controlling central network scheduler. Using tokens as a global tool to enforce traffic limiting and policies, the scheduler is able to proactively control the flow of traffic as well as respond to application requests. "Knowledge" and predictive analysis in networks is a growing trend in research which we intend to leverage in this system.
We investigate the potential in such a new form of interactive congestion control and analyze it against to state-of-the-art solutions. Our tool of choice to model our system is Mininet, a rapid prototyping emulator for datacenter networks. We will compare our "TCPToken" system against established TCP-congestion systems such as Hedera and DCTCP. Based on the findings and measurement results, we will reassess the feasibility and prospect of a token-based, predictive congestion control algorithm.
\end{abstract}

\keywords{TCP, Congestion Control, SDN, Data center}

\maketitle

\input{intro}

\input{relatedworks}
\input{design}
%\input{motivation}
\input{implementation}
\input{eval}
\input{timeline}

\bibliographystyle{ACM-Reference-Format}
\bibliography{paper}

\end{document}