### 2017-11-21 Documents/Proposal/timeline.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Timeline}
\label{sec:timeline}
\begin{itemize}
\item \underline{October 15th:} Finalize draft and related work / background section.
\item \underline{ October 22nd:} Iterate over literature and thoroughly analyze related work. Set up an initial prototyping environment in Mininet on which each team member can work on.
\item \underline{October 30th:} Finalize the initial simple design and explore naive congestion modelling in Mininet. Integrate existing Mininet emulations of Hedera and DCTCP.
\item \underline{November 12th:} Develop initial naive TCPTokens prototype involving controller and end-host design.
\item \underline{November 19th:} Develop intelligent centralised congestion control heuristic.
\item \underline{November 26th:} Run evaluations and microbenchmarks.
\item \underline{December 3rd:} Iterate and improve the scheduler, explore topology and bandwidth constraints. Test at scale.
\item \underline{December 10th:} Write report and prepare presentation.
\end{itemize}