### 2017-12-09 Documents/Report/eval.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Evaluation}
\label{sec:evaluation}

In the absence of a commercial data center, the implementation of our network 
design is going to be done on top of mininet with simulated FatTree topologies 
of varying size. We stress test with iPerf and simulate data center traffic 
using tcpreplay and packet traces provided by~\cite{traffic} .

To evaluate the general effectiveness of our system, we plan to measure against 
existing centralized as well as decentralized solutions.
The centralized design will be based on Hedera~\cite{hedera}, a common and 
influential datacenter scheduler. The decentralized congestion control 
mechanism will be DCTCP~\cite{dctcp}, a state-of-the-art TCP congestion 
algorithm.

Since we are primarily concerned about reducing the latency and packet drops 
while keeping utilisation at maximum, the measurements to get a good insight 
into how well the design can perform are as follows:
\begin{enumerate}

\item Latency: We are aiming for a low latency network which means that 99th 
percentile latency in the network across all flows should as low as possible. 
Latency should be measured when all hosts are sending packets at the maximum 
limit and also with random traffic patterns. There should be low latency even 
during sudden bursts as the transmission rate is limited by a base value.

\item Packet drop rate: Ideally this metric will approach zero, as the 
objective of Iroko is to minimize packet loss.

\item Fairness: Fairness can be tested by introducing a new host to a 
completely saturated network and increasing the transmission rate to see if all 
the hosts gets fair share of the total bandwidth. We are assuming all flows 
should be at equal priority. Differential priority is out of scope.

\item Responsiveness: This metric depends on the predictive power and 
efficiency of Iroko. It may be interesting to analyze the response time of the 
central scheduler compared to decentralized DCTCP and TCP. In addition, it is 
valuable to identify at what data center and flow size the central computing 
node of the network may become a bottleneck.

\item Network utilization: We will measure the used bandwidth and the total 
load on the hosts. These metrics can be measured by varying the load from each 
host. Load to utilization ratio should be 1 until the saturation point and 
after that utilization should stay stable at the maximum bisection bandwidth.
It is also important to check that there is no starvation happening in any 
host. This can be measured by simulating random traffic patterns and plotting 
residual bandwidth and excess load. We expect overall utilization to be lower 
compared to Hedera and DCTCP. Although Iroko will maximize goodput, we do not 
measure it on grounds of simplicity.

\end{enumerate}

As baseline we will also compare to the default TCP congestion algorithm to 
estimate how our scheduler fares against the default case. As our initial 
simple design is very constrained in its ability to route and control traffic 
we expect to achieve less overall optimality against DCTCP and Hedera, but 
still gain a substantial advantage over common TCP.