### 2017-12-10 Documents/Report/conclusion.tex
Added discussion, future work and conclusion. Sparse thoughts for first two.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}