### 2017-12-10 Documents/Report/Iroko_Report.tex
more changes to intro

\documentclass[sigconf]{acmart}

\usepackage{booktabs} % For formal tables
\usepackage[normalem]{ulem}
% Copyright
\setcopyright{none}

\input{annotations}

\begin{document}

\title{Iroko: Predictive Datacenter Congestion Control}

\author{Anand Jayarajan}
\affiliation{%
  \institution{University of British Columbia}
}
\email{anandj@cs.ubc.ca}

\author{Michael Przystupa}
\affiliation{%
  \institution{University of British Columbia}
}
\email{michael.przystupa@gmail.com}

\author{Robert Reiss}
\affiliation{%
  \institution{University of British Columbia}
}
\email{rreiss@cs.ubc.ca}

\author{Fabian Ruffy}
\affiliation{%
  \institution{University of British Columbia}
}
\email{fruffy@cs.ubc.ca}

\begin{abstract}
Datacenter networks have become a hotbed of network research in the past 
decade. Many works have been published on improving bisection bandwidth, cost, 
and latency. As a core contribution of this research, centralized scheduling 
has entered the stage as a dominant strategy to manage flows. Benefitting from 
a global view and full control over the network, centralized schedulers are 
able to enforce fine-grained traffic control, frequently achieving near-optimal 
bandwidth optimization. However, modern congestion control algorithms and 
schedulers are still fundamentally reactive. Many algorithms are designed to 
only respond to packet loss or significant increase in latency, not to actively 
prevent it.
In this project, we plan to explore the opportunities of employing a preemptive 
and tightly controlling central network scheduler. By enforcing traffic limits 
and admission control, the scheduler is able to proactively steer the flow of 
traffic. The notion of "knowledge" and predictive analysis in networks is a 
growing trend in research, which we intend to leverage in this system.

We investigate the potential in such a new form of interactive congestion 
control and analyze it against state-of-the-art solutions. Our tool of choice 
to model our system is Mininet, a rapid prototyping emulator for datacenter 
networks. We will compare our "Iroko" system against established TCP-congestion 
systems such as Hedera and DCTCP. Based on the findings, measurement results, 
and effectiveness of our reinforcement learning technique, we will reassess the 
feasibility and prospects of a predictive congestion control algorithm.
\end{abstract}

\keywords{TCP, Congestion Control, SDN, Datacenter}

\maketitle

\input{intro}
%\input{motivation}
\input{design}

\input{implementation}
\input{eval}
\input{discussion}
\input{future_work}
\input{conclusion}

\bibliographystyle{ACM-Reference-Format}
\bibliography{Iroko_Report}

\end{document}