### 2017-12-10 Documents/Report/related_work.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}
\label{sec:related}
Network traffic congestion control and fairness has been an active topic of 
research for several decades and exhibits a vast range of work. We draw the 
majority of our inspiration from the following four categories.
\subsection{End-to-End TCP optimization}
\subsection{Centralized Scheduling and Congestion Control}
\subsection{Network Admission-Control}
\subsection{Datacenter Predictability and Learning}