### 2017-12-12 Documents/Report/conclusion.tex
Edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
\label{sec:conclusion}

Iroko represents a first attempt at using machine learning
to predict datacenter traffic. The endeavour relies crucially
on datacenter traffic having enough structure to in fact allow
some level of prediction. Yet even for largely random traffic,
Iroko showed decent results, often besting baseline solutions.
For more predictable traffic, Iroko can compete with "best-in-class"
solutions, and at times shows the strongest performance of all the
tested solutions.
Iroko is only the opening salvo, but the attempt has merit. Iroko
struggles with reducing drop rate, and some random traffic
patterns. With more fine-tuned
prediction, and better rate-limiting algorithms, Iroko can likely
overcome these challenges. 
More work must be done before claims of superiority to all
incumbent solutions can be made, but there are reasons to be
optimistic.

Machine learning is perhaps the king of "black-box" solutions, 
and its efficacy is undeniable, if sometimes misleading. While
the standoff between so called "white-box" and "black-box" solutions is 
likely to continue for the foreseeable future, Iroko shows
that "black-box" solutions do have a role to play. It might be the case
that neither comes to rule, but rather future SDN's will 
employ both techniques to build ever more efficient solutions.
It may well turn out that the future is grey.